---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# icstest

<!-- badges: start -->
<!-- badges: end -->

The goal of **icstest** is to provide methods for testing informative cluster size (ICS) in cluster randomized trials. Informative cluster size occurs when treatment effects depend on the size of clusters, which can bias standard analyses that ignore this relationship.

## Installation

You can install the development version of icstest from GitHub with:

``` r
devtools::install_github("Zhe-Chen-1999/icstest")
```

## Overview

The package implements three approaches for testing informative cluster size:

- **Model-Assisted Test**: Test coefficient for treatment in an ordinary least squares (OLS) regression of weighted cluster means, where weights are proportional to deviations of each cluster size from the average cluster size
- **Randomization-Based Test**: Permute the cluster-level treatments and compute the model-assisted test statistic for each permutation. The p-value is the proportion of permuted test statistics that are as or more extreme than the observed test statistic
- **Model-Based Test**: Test coefficients for interactions between treatment and cluster size in a generalized estimating equation (GEE) outcome model

## Example

This is a basic example which shows you how to test for informative cluster size:

```{r example}
library(icstest)

# Generate example data with informative cluster sizes
set.seed(123)
n_clusters <- 100
cluster_sizes <- sample(20:80, n_clusters, replace = TRUE)
cluster_size <- rep(cluster_sizes, cluster_sizes)
n <- sum(cluster_sizes)
cluster_id <- rep(1:n_clusters, cluster_sizes)
Z <- rep(rbinom(n_clusters, 1, 0.5), cluster_sizes)

# Generate outcomes with cluster size-dependent treatment effects
Y0 <- rnorm(n) + rep(rnorm(n_clusters), cluster_sizes)
tau <- rnorm(n) + rep(1.5 * (cluster_sizes > 50), cluster_sizes)
Y <- Y0 + tau * Z

# Run model-assisted test
ma_result <- model_assisted_test(Y, Z, cluster_id)

# Randomization-Based Test
rand_result <- randomization_test(Y, Z, cluster_id, n_perms = 5000)

# Model-Based Test
data <- data.frame(
  Y = Y, 
  Z = Z, 
  cluster_size = cluster_size,
  cluster_id = cluster_id
)

formula <- Y ~ Z * cluster_size # Test linear interaction
mb_result <- model_based_test(formula, "Z:cluster_size", data, cluster_id)
```
